---
title: "CBD COP Analysis"
author: "Jeff Cegan"
date: "`r Sys.Date()`"
output: html_document
editor_options:
  chunk_output_type: console
---

<br>

```{r global-options, include=TRUE, echo=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=6, fig.align = 'left', fig.path='knitr_figs/',
                          include=FALSE,echo=FALSE, warning=FALSE, message=FALSE)
#include chunk, echo code, don't print messages or warnings
```

```{r load packages,echo=FALSE}
pkgs = c('lubridate','downloadthis','plotly','skimr','lemon',"tidyverse")
inst = lapply(pkgs, library, character.only = TRUE) # load them
knit_print.data.frame <- lemon_print
```

```{r read input files}
loc <- read.csv('input/cbd_location.csv', stringsAsFactors=TRUE)
  loc <- loc %>% mutate(Year=year(ymd(loc$Date))) #add year column
  loc <- loc %>% mutate(Month=month(ymd(loc$Date))) #add year column 
  loc <- loc %>% mutate(Date=date(paste0(Year,'-',Month,"-01"))) #add year column
att <- read.csv('input/cbd_attendance.csv', stringsAsFactors=TRUE)
fil <- read.csv('input/filters.csv', stringsAsFactors=TRUE) %>% rename(CountryCode = Country.code, StateTerritory = State.Territory)
```

```{r}
#clean attendance dataset
#converted all 'n/a' to NA manually in excel
#redundant information in state territory (deleted from CBD attendance)
#pivot longer
attl <- att %>% select(!Party_birth) %>% 
  pivot_longer(cols=starts_with('COP'),names_to = 'Conference', values_to = 'Delegates')
```

```{r}
#checking merges
  length(unique(attl$CountryCode))
  length(unique(fil$CountryCode))
inner<- inner_join(attl,fil, by = 'CountryCode')
  length(unique(inner$CountryCode))

#join attendance and filter tables
df <- left_join(attl, fil, by = 'CountryCode')
  length(unique(df$CountryCode))

#join location table
df <- left_join(df, loc, by = 'Conference')
```

<br>

```{r}
#create custom theme for graphs
theme_cegan <- function(){ 
    font <- "Calibri"   #assign font family up front
    
    theme_bw() %+replace%    #replace elements we want to change
    
    theme(
      
      #grid elements
      # panel.grid.major = element_blank(),    #strip major gridlines
      # panel.grid.minor = element_blank(),    #strip minor gridlines
      # axis.ticks = element_blank(),          #strip axis ticks
      
      #text elements
      plot.title = element_text(             #title
                   family = font,            #set font family
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,               #center align
                   vjust = 2),               #raise slightly
      
      plot.subtitle = element_text(          #subtitle
                   family = font,            #font family
                   size = 14),               #font size
      
      plot.caption = element_text(           #caption
                   family = font,            #font family
                   size = 9,                 #font size
                   hjust = 1),               #right align
      
      axis.title = element_text(             #axis titles
                   family = font,            #font family
                   size = 16),               #font size
      
      axis.text = element_text(              #axis text
                   family = font,            #axis family
                   face = 'bold',            #bold typeface
                   size = 14),               #font size
      
      axis.text.x = element_text(            #margin for axis text
                    margin=margin(5, b = 10)),

      axis.title.y = element_text(           #margin for axis text
                    angle = 90,
                    margin = margin(t = 0, r = 10, b = 0, l = 0)),
      
      # axis.title.y = element_text(           #margin for axis text
      #               margin = margin(t = 0, r = 10, b = 0, l = 0)),
      #       
      legend.text = element_text(
                    family = font,          #axis family
                    size=14)                #font size
    )
}
```

<br>

<h2 align="center">

Summary Tables

</h2>

<br>

#### All 196 countries from the CBD attendance table have a filter counterpart (LDC/SIDS assignment)

```{r}
att %>% filter(!CountryCode %in% 
  unique(inner$CountryCode)) %>% 
  distinct(CountryCode) %>% 
  rename('Country Code' = CountryCode)
```

<br>

#### Countries from the filter table (LDC/SIDS information) without CBD attendance

<br>

```{r, include=TRUE, echo=F, render=lemon_print}
#tinas_selection <- c('AND','PRK','SSD','NIU','PSE') 
  
fil %>% select(CountryCode, StateTerritory) %>% 
  filter(!CountryCode %in% unique(inner$CountryCode)) %>% 
  distinct(CountryCode, StateTerritory) %>% 
  rename('State Territory' = StateTerritory,'Country Code' = CountryCode) 
#filter(CountryCode %in% tinas_selection)
```

<br>

#### LDC vs. SIDS (Total Countries)

<br>

```{r}
#View(df)
begin<-nrow(df)
df %>% filter(is.na(LDC) | is.na(SIDS)) #14 EUE entries missing label


nrow(df)
df <- df %>% 
  na.omit() %>% 
  mutate(Status = ifelse(LDC == 1 & SIDS == 0, "LDC",
                                     ifelse(LDC == 0 & SIDS == 1, "SIDS",
                                            ifelse(LDC == 1 & SIDS == 1, "Both", 'Neither'))))
end <- nrow(df) 
begin - end #dropped the 275 NA attendance
length(unique(df$CountryCode)) #196 unique countries 
```

```{r, include=TRUE, echo=F, render=lemon_print}
#count number of countries in status category for table
zdf_table <- df %>%
  select(Status, CountryCode) %>%
  group_by(Status) %>%
  distinct() %>%
  count(Status)

zdf_table <- data.frame(zdf_table)
zdf_table
```

<br>

<h2 align="center">

Median Graphs

</h2>

<br>

<br>

#### 4 Categories: LDC vs. SIDS (Median Delegates)

<br>

```{r}
## 4 CATEGORIES ##
#aggregate by status for graph
zdf <- df %>% 
  group_by(Status,Date) %>% 
  summarize(Median_Delegates = median(Delegates),Mean_Delegates = mean(Delegates), n = n())

## 3 CATEGORIES ##
#Both to move into LDC and SIDS counts as seperate entries
nrow(df)
  #subset of Both LDC and SIDS rows
  zdf_noboth_1 <- df %>% filter(Status == 'Both')
    nrow(zdf_noboth_1)
  zdf_noboth_1a <- zdf_noboth_1 %>% mutate(Status = 'LDC')
  zdf_noboth_1b <- zdf_noboth_1 %>% mutate(Status = 'SIDS')
  #delete LDC and SIDS rows
  zdf_noboth_2 <- df %>% filter(Status != 'Both')  
    nrow(zdf_noboth_2)    
  #append subset
  zdf_noboth_3 <- bind_rows(zdf_noboth_1a,zdf_noboth_1b,zdf_noboth_2)
#check
  nrow(zdf_noboth_3) == nrow(df) + nrow(zdf_noboth_1)
 
zdf_noboth <- zdf_noboth_3 %>% 
  group_by(Status,Date) %>% 
  summarize(Median_Delegates = median(Delegates),Mean_Delegates = mean(Delegates), n = n())
```

```{r, include=TRUE, echo=F}
## GRAPH 4 CATEGORIES ##
#order factors for legend
zdf$Status <- factor(zdf$Status, 
      levels=c("Neither","LDC",'SIDS','Both'),
      labels=c("Neither","LDC",'SIDS','Both'))

#Graph
fig <- zdf %>% 
  na.omit() %>% 
  ggplot(aes(Date,Median_Delegates)) + 
  geom_line(aes(color=Status, linetype=Status),lwd=.8) +
  geom_point() +
  theme_cegan() +
  labs(x="Date", y='Median Delegates') + 
  scale_color_discrete(name="") + 
  scale_linetype_discrete(name = "") + 
  scale_x_date(date_breaks = "5 years", 
               date_labels = "%Y",
               limits = as.Date(c("1992-01-01","2018-01-01"))) 

fig <- fig + theme(legend.title = element_blank(),
            legend.position = c(0.15, 0.80),
            legend.background = element_rect(size=0.4, linetype = 'solid', color='black')) 

fig

fig %>% download_this(output_name = "CBD_4Cat_Median",
    output_extension = ".png",
    button_label = "Download Graph",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save")
```

------------------------------------------------------------------------

<br>

#### 3 Categories: LDC vs. SIDS (Median Delegates)

<br>

```{r, include=TRUE, echo=F}
## GRAPH 3 CATEGORIES ##
#order factors for legend
zdf_noboth$Status <- factor(zdf_noboth$Status, 
      levels=c("Neither","LDC",'SIDS'),
      labels=c("Neither","LDC",'SIDS'))

#Graph
fig <- zdf_noboth %>% 
  na.omit() %>% 
  ggplot(aes(Date,Median_Delegates)) + 
  geom_line(aes(color=Status, linetype=Status),lwd=.8) +
  geom_point() +
  theme_cegan() +
  labs(x="Date", y='Median Delegates') + 
  scale_color_discrete(name="") + 
  scale_linetype_discrete(name = "") + 
  scale_x_date(date_breaks = "5 years", 
               date_labels = "%Y",
               limits = as.Date(c("1992-01-01","2018-01-01"))) 

fig <- fig + theme(legend.title = element_blank(),
            legend.position = c(0.15, 0.80),
            legend.background = element_rect(size=0.4, linetype = 'solid', color='black')) 

fig

fig %>% download_this(output_name = "CBD_3Cat_Median",
    output_extension = ".png",
    button_label = "Download Graph",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save")
```

------------------------------------------------------------------------

<br>

<h2 align="center">

Mean Graphs

</h2>

<br>

#### 4 Categories: LDC vs. SIDS (Mean Delegates)

<br>

```{r, include=TRUE, echo=F}
## GRAPH 4 CATEGORIES ##
#order factors for legend
zdf$Status <- factor(zdf$Status, 
      levels=c("Neither","LDC",'SIDS','Both'),
      labels=c("Neither","LDC",'SIDS','Both'))

#Graph
fig <- zdf %>% 
  na.omit() %>% 
  ggplot(aes(Date,Mean_Delegates)) + 
  geom_line(aes(color=Status, linetype=Status),lwd=.8) +
  geom_point() +
  theme_cegan() +
  labs(x="Date", y='Mean Delegates') + 
  scale_color_discrete(name="") + 
  scale_linetype_discrete(name = "") + 
  scale_x_date(date_breaks = "5 years", 
               date_labels = "%Y",
               limits = as.Date(c("1992-01-01","2018-01-01"))) 

fig <- fig + theme(legend.title = element_blank(),
            legend.position = c(0.15, 0.80),
            legend.background = element_rect(size=0.4, linetype = 'solid', color='black')) 

fig

fig %>% download_this(output_name = "CBD_4Cat_Mean",
    output_extension = ".png",
    button_label = "Download Graph",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save")
```

------------------------------------------------------------------------

<br>

#### 3 Categories: LDC vs. SIDS (Mean Delegates)

<br>

```{r, include=TRUE, echo=F}
## GRAPH 3 CATEGORIES ##
#order factors for legend
zdf_noboth$Status <- factor(zdf_noboth$Status, 
      levels=c("Neither","LDC",'SIDS'),
      labels=c("Neither","LDC",'SIDS'))

#Graph
fig <- zdf_noboth %>% 
  na.omit() %>% 
  ggplot(aes(Date,Mean_Delegates)) + 
  geom_line(aes(color=Status, linetype=Status),lwd=.8) +
  geom_point() +
  theme_cegan() +
  labs(x="Date", y='Mean Delegates') + 
  scale_color_discrete(name="") + 
  scale_linetype_discrete(name = "") + 
  scale_x_date(date_breaks = "5 years", 
               date_labels = "%Y",
               limits = as.Date(c("1992-01-01","2018-01-01"))) 

fig <- fig + theme(legend.title = element_blank(),
            legend.position = c(0.15, 0.80),
            legend.background = element_rect(size=0.4, linetype = 'solid', color='black')) 

fig

fig %>% download_this(output_name = "CBD_3Cat_Mean",
    output_extension = ".png",
    button_label = "Download Graph",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save")
```

------------------------------------------------------------------------

<br>

#### 3 Categories: LDC vs. SIDS (Mean Delegates) with Labels

<br>

```{r}
#merge back loc cop information
nrow(zdf_noboth)
zdf_noboth_l <- inner_join(zdf_noboth,loc,by='Date')
zdf_noboth_l <- zdf_noboth_l %>% 
  mutate(Conference = ifelse(Status != 'Neither'," ",as.character(Conference)))
nrow(zdf_noboth_l) == nrow(zdf_noboth)
```

```{r, include=TRUE, echo=F}
## GRAPH 3 CATEGORIES ##


#Graph
fig <- zdf_noboth_l %>% 
  na.omit() %>% 
  ggplot(aes(Date,Mean_Delegates)) + 
  geom_line(aes(color=Status, linetype=Status),lwd=.8) +
  geom_point() +
  geom_text(aes(label = Conference),nudge_y=1,check_overlap=T) +
  theme_cegan() +
  labs(x="Date", y='Mean Delegates') + 
  scale_color_discrete(name="") + 
  scale_linetype_discrete(name = "") + 
  scale_x_date(date_breaks = "5 years", 
               date_labels = "%Y",
               limits = as.Date(c("1992-01-01","2018-01-01"))) 

fig <- fig + theme(legend.title = element_blank(),
            legend.position = c(0.15, 0.80),
            legend.background = element_rect(size=0.4, linetype = 'solid', color='black')) 

fig 

fig %>% download_this(output_name = "CBD_3Cat_Mean_Labels",
    output_extension = ".png",
    button_label = "Download Graph",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save")
```

------------------------------------------------------------------------

<br> <br>
