---
title: "UNFCCC COP Analysis"
author: "Jeff Cegan"
date: "`r Sys.Date()`"
output: html_document
editor_options:
  chunk_output_type: console
---

<br>

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=6, fig.align = 'left', fig.path='knitr_figs/',
                          include=FALSE,echo=TRUE, warning=FALSE, message=FALSE)
#include chunk, echo code, don't print messages or warnings
```

```{r load packages}
pkgs = c('downloadthis','extrafont','plotly','here','skimr','janitor','plotly','lemon',"tidyverse")
inst = lapply(pkgs, library, character.only = TRUE) # load them
knit_print.data.frame <- lemon_print
```

```{r read input files}
att <- read.csv('input/attendance.csv', stringsAsFactors=TRUE)
fil <- read.csv('input/filters.csv', stringsAsFactors=TRUE) %>% rename(CountryCode = Country.code, StateTerritory = State.Territory)
```

```{r}
#checking merges
  length(unique(att$CountryCode))
  length(unique(fil$CountryCode))
inner<- inner_join(att,fil, by = 'CountryCode')
  length(unique(inner$CountryCode))
df <- left_join(att, fil, by = 'CountryCode')
  length(unique(df$CountryCode))

#cop list for graphics
Conference <- paste0("COP ",1:21)
```

```{r}
#create custom theme for graphs
theme_cegan <- function(){ 
    font <- "Calibri"   #assign font family up front
    
    theme_bw() %+replace%    #replace elements we want to change
    
    theme(
      
      #grid elements
      # panel.grid.major = element_blank(),    #strip major gridlines
      # panel.grid.minor = element_blank(),    #strip minor gridlines
      # axis.ticks = element_blank(),          #strip axis ticks
      
      #text elements
      plot.title = element_text(             #title
                   family = font,            #set font family
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,               #center align
                   vjust = 2),               #raise slightly
      
      plot.subtitle = element_text(          #subtitle
                   family = font,            #font family
                   size = 14),               #font size
      
      plot.caption = element_text(           #caption
                   family = font,            #font family
                   size = 9,                 #font size
                   hjust = 1),               #right align
      
      axis.title = element_text(             #axis titles
                   family = font,            #font family
                   size = 16),               #font size
      
      axis.text = element_text(              #axis text
                   family = font,            #axis family
                   face = 'bold',            #bold typeface
                   size = 14),               #font size
      
      axis.text.x = element_text(            #margin for axis text
                    margin=margin(5, b = 10)),

      axis.title.y = element_text(           #margin for axis text
                    angle = 90,
                    margin = margin(t = 0, r = 10, b = 0, l = 0)),
      
      # axis.title.y = element_text(           #margin for axis text
      #               margin = margin(t = 0, r = 10, b = 0, l = 0)),
      #       
      legend.text = element_text(
                    family = font,          #axis family
                    size=14)                #font size
    )
}
```

<br>

#### Countries from the attendance table without a filter counterpart (LDC/SIDS assignment)

SCG did not have a counterpart. SCG was both Serbia and Montenegro prior to 2006. Now SCG is divided into two countries, SRB and MNE. For this analysis, SCG attendance was assigned to SRB for the 1995-2005 time period.

```{r}
#SCG was Serbia and Montenegro when they still were one country prior to 2006. Now SCG is divided into SRB and MNE. Assign SCG attendance to SRB
#SCG and SRB don't overlap in time '95-'05; '96-'15 respectively
#Replace SCG label with SRB
att <- att %>% mutate(CountryCode = replace(CountryCode, CountryCode == 'SCG','SRB'))
```

```{r}
att %>% filter(!CountryCode %in% 
  unique(inner$CountryCode)) %>% 
  distinct(CountryCode) %>% 
  rename('Country Code' = CountryCode)
```

<br>

#### Countries (Tina's subset) from the filter table (LDC/SIDS information) without COP attendance

<br>

```{r, include=TRUE, echo=F, render=lemon_print}
tinas_selection <- c('AND','PRK','SSD','NIU','PSE') 
  
fil %>% select(CountryCode, StateTerritory) %>% 
  filter(!CountryCode %in% unique(inner$CountryCode) & CountryCode %in% tinas_selection) %>% 
  distinct(CountryCode, StateTerritory) %>% 
  rename('State Territory' = StateTerritory,'Country Code' = CountryCode) 
#filter(CountryCode %in% tinas_selection)
```

<br> <br>

#### Time Series: LDC vs. SIDS (Mean Delegates)

<br>

```{r}
nrow(df)
df <- df %>% 
  na.omit() %>% 
  mutate(Status = ifelse(LDC == 1 & SIDS == 0, "LDC",
                                     ifelse(LDC == 0 & SIDS == 1, "SIDS",
                                            ifelse(LDC == 1 & SIDS == 1, "Both", 'Neither'))))
nrow(df) #dropped the 11 unmatched SCG rows
length(unique(df$CountryCode)) #191 unique countries 
```

```{r, include=TRUE, echo=F, render=lemon_print}
#count number of countries in status category for table
zdf_table <- df %>%
  select(Status, CountryCode) %>%
  group_by(Status) %>%
  distinct() %>%
  count(Status)

zdf_table <- data.frame(zdf_table)
zdf_table
```

<br>

```{r}
## 4 CATEGORIES ##
#aggregate by status for graph
zdf <- df %>% 
  group_by(Status,Year) %>% 
  summarize(Median_Delegates = median(Delegates),Mean_Delegates = mean(Delegates), n = n())

## 3 CATEGORIES ##
#Both to move into LDC and SIDS counts as seperate entries
nrow(df)
  #subset of Both LDC and SIDS rows
  zdf_noboth_1 <- df %>% filter(Status == 'Both')
    nrow(zdf_noboth_1)
  zdf_noboth_1a <- zdf_noboth_1 %>% mutate(Status = 'LDC')
  zdf_noboth_1b <- zdf_noboth_1 %>% mutate(Status = 'SIDS')
  #delete LDC and SIDS rows
  zdf_noboth_2 <- df %>% filter(Status != 'Both')  
    nrow(zdf_noboth_2)    
  #append subset
  zdf_noboth_3 <- bind_rows(zdf_noboth_1a,zdf_noboth_1b,zdf_noboth_2)
#check
  nrow(zdf_noboth_3) == nrow(df) + nrow(zdf_noboth_1)
 
zdf_noboth <- zdf_noboth_3 %>% 
  group_by(Status,Year) %>% 
  summarize(Median_Delegates = median(Delegates),Mean_Delegates = mean(Delegates), n = n())
```

```{r, include=TRUE, echo=F}
## GRAPH 4 CATEGORIES ##
#order factors for legend
zdf$Status <- factor(zdf$Status, 
      levels=c("Neither","LDC",'SIDS','Both'),
      labels=c("Neither","LDC",'SIDS','Both'))

#Graph
fig <- zdf %>% 
  na.omit() %>% 
  ggplot(aes(Year,Median_Delegates)) + 
  geom_line(aes(color=Status, linetype=Status),lwd=.8) +
  geom_point() +
  theme_cegan() +
  labs(x="Year", y='Median Delegates') + 
  scale_color_discrete(name="") + 
  scale_linetype_discrete(name = "")

fig <- fig + theme(legend.title = element_blank(),
            legend.position = c(0.15, 0.80),
            legend.background = element_rect(size=0.4, linetype = 'solid', color='black')) 

# scale_linetype_manual(values = c("Neither" = 'solid',"LDC" = 'dotted','SIDS' = 'solid','Both' = 'dotted')) +  
#   guides(linetype = FALSE)

fig

fig %>% download_this(output_name = "UNFCCC_4Cat_Median",
    output_extension = ".png",
    button_label = "Download Graph",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save")
```

<br>

```{r, include=TRUE, echo=F}
## GRAPH 3 CATEGORIES ##
#order factors for legend
zdf_noboth$Status <- factor(zdf_noboth$Status, 
      levels=c("Neither","LDC",'SIDS'),
      labels=c("Neither","LDC",'SIDS'))

#Graph
fig <- zdf_noboth %>% 
  na.omit() %>% 
  ggplot(aes(Year,Median_Delegates)) + 
  geom_line(aes(color=Status, linetype=Status),lwd=.8) +
  geom_point() +
  theme_cegan() +
  labs(x="Year", y='Median Delegates') + 
  scale_color_discrete(name="") + 
  scale_linetype_discrete(name = "")

fig <- fig + theme(legend.title = element_blank(),
            legend.position = c(0.15, 0.80),
            legend.background = element_rect(size=0.4, linetype = 'solid', color='black')) 

fig

fig %>% download_this(output_name = "UNFCCC_3Cat_Median",
    output_extension = ".png",
    button_label = "Download Graph",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save")
```

<br> <br>

#### Time Series: All 191 Attending Countries (Mean Delegates)

<br>

```{r, aggregate LDC, SIDS, Non-SIDS}
#create LDC or SIDS column
rowstart <- nrow(df)
df <- df %>% mutate(LDC_SIDS = ifelse(LDC == 1 | SIDS == 1, "LDC or SIDS","Non LDC/SIDS"))
rowend <- nrow(df)
rowstart == rowend

#aggregate by status
adf <- df %>% group_by(LDC_SIDS,Year) %>% summarize(Median_Delegates = median(Delegates),Mean_Delegates = mean(Delegates), n = n())
```

```{r}
#order factors for legend
adf$LDC_SIDS <- factor(adf$LDC_SIDS, 
      levels=c("Non LDC/SIDS","LDC or SIDS"),
      labels=c("Non LDC/SIDS","LDC or SIDS"))
```

```{r, include=TRUE, echo=F}
adf <- adf %>% na.omit() %>% mutate(Conference = c(Conference))

fig <- adf %>% 
  na.omit() %>% 
  ggplot(aes(Year,Mean_Delegates,label=Conference)) + 
  geom_line(aes(color=LDC_SIDS, linetype=LDC_SIDS),lwd=1) +
  geom_point() +
  theme_cegan() +
  labs(x="Year", y='Mean Delegates') + 
  scale_color_discrete(name="") + 
  scale_linetype_discrete(name = "")

fig # %>% ggplotly(tooltip = c('Conference','Mean_Delegates'))

fig %>% download_this(output_name = "UNFCCC_01_MeanDelegates",
    output_extension = ".png",
    button_label = "Download Graph",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save")
```

<br> <br>

#### Time Series: All 191 Attending Countries (Median Delegates)

<br>

```{r, include=TRUE, echo=F}
adf <- adf %>% na.omit() %>% mutate(Conference = c(Conference))

fig <- adf %>% 
  na.omit() %>% 
  ggplot(aes(Year,Median_Delegates,label=Conference)) + 
  geom_line(aes(color=LDC_SIDS, linetype=LDC_SIDS),lwd=1) +
  geom_point() +
  theme_cegan() +
  labs(x="Year", y='Median Delegates') + 
  scale_color_discrete(name="") + 
  scale_linetype_discrete(name = "")

fig # %>% ggplotly(tooltip = c('Conference','Median_Delegates'))

fig %>% download_this(output_name = "UNFCCC_01_MedianDelegates",
    output_extension = ".png",
    button_label = "Download Graph",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save")
```

<br> <br>

#### Multiplier of Non-LDC/SIDS vs. LDC/SIDS Countries

<br>

$$\frac{mean(Delegates_{LDC/SIDS})_{year_i}}{mean(Delegates_{non-LDC/SIDS})_{year_i}}
$$

<br>

```{r}
#LDC_SIDS column to factor prior to pivot wider
#Calculate percentage increase of LDC/non-LDC for graph
adf$LDC_SIDS <- as.factor(gsub("/","",gsub(" ","",adf$LDC_SIDS)))

adf_perc <- adf %>% select(LDC_SIDS,Year,Mean_Delegates) %>% 
  pivot_wider(names_from = LDC_SIDS, values_from = Mean_Delegates) %>% 
  mutate(Multiplier = NonLDCSIDS/LDCorSIDS)

adf_perc
# zdf

```

### Non-LDC SIDS: Share at the Table

<br>

#### [Interactive Plot:]{style="color: #795695;"}

```{r, include=TRUE, echo=F}
#Bar graph LDC/non-LDC percentage over time
fig<-adf_perc %>%  
  ggplot(aes(x=Year, y=Multiplier,label=Conference)) + 
  geom_bar(stat="identity", fill='steel blue') +
  labs(x="Year", y='Multiplier') +
  theme_cegan()
  # scale_y_continuous(labels = scales::percent_format(scale = 1))

pfig <- fig + geom_bar(stat="identity", fill='#795695') 
pfig %>% ggplotly(tooltip = c('Multiplier','Conference'), height= 500, width = 670)
```

<br>

#### [Plot for Paper (higher-quality download):]{style="color: #4682B4;"}

```{r, include=TRUE, echo=F, fig.height=5, fig.width=7}
fig
```

```{r, include=TRUE, echo=F}
fig %>% download_this(output_name = "UNFCCC_02_Multiplier",
    output_extension = ".png",
    button_label = "Download Graph",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save")
```

<br> <br>

#### Time Series: Top 10 of 105 Attending Non-LDC/SIDS Countries

<br>

```{r, various non-ldc/sids aggregations}
#filter by status aggregate by country
tempdf <- df %>% filter(LDC_SIDS == "Non LDC/SIDS") %>% group_by(StateTerritory) %>% summarize(Mean_Delegates = mean(Delegates), n = n()) %>% arrange(desc(Mean_Delegates))

topcountries <- tempdf[1:10,'StateTerritory']
bottomcountries <- tempdf[11:nrow(tempdf),'StateTerritory']

bdf_bottomcountries_all <- df %>% 
  filter(LDC_SIDS == "Non LDC/SIDS", StateTerritory %in% bottomcountries$StateTerritory) %>% 
  group_by(StateTerritory) %>% 
  summarize(Mean_Delegates = mean(Delegates), n = n())

bdf_bottomcountries <- df %>% 
  filter(LDC_SIDS == "Non LDC/SIDS", StateTerritory %in% bottomcountries$StateTerritory) %>% 
  group_by(Year) %>% 
  summarize(Mean_Delegates = mean(Delegates), n = n()) %>% 
  mutate(StateTerritory = as.factor('Remaining Average'))

bdf_topcountries <- df %>% 
  filter(LDC_SIDS == "Non LDC/SIDS") %>% 
  group_by(StateTerritory,Year) %>% 
  summarize(Mean_Delegates = mean(Delegates), n = n()) %>% 
  filter(StateTerritory %in% topcountries$StateTerritory)

#append top and bottom for same graph
bdf <- bind_rows(bdf_topcountries,bdf_bottomcountries)

bdf_avg <- bdf %>% 
  group_by(StateTerritory) %>% 
  summarise(Mean_Delegates = mean(Mean_Delegates))

#append top and bottom for table
bdf_avg_all <- rbind(
  bdf_topcountries %>% 
    group_by(StateTerritory) %>% 
    summarize(Mean_Delegates =mean(Mean_Delegates)),
  bdf_bottomcountries_all %>% 
    group_by(StateTerritory) %>% 
    summarize(Mean_Delegates =mean(Mean_Delegates))
  )
```

```{r, include=TRUE, echo=F}

#legend labels in order, modified names
  names <- c("Brazil","France","USA","Japan","Canada","China","Korea","Denmark","Germany","Indonesia","Other Non-LDC/SIDS")
  
#order values for labels  
  m_values <- bdf_avg %>% select(Mean_Delegates) %>% arrange(desc(Mean_Delegates))
  values = paste0("(",as.character(as.integer(round(m_values$Mean_Delegates,0))),")")
  legend_labels <- paste(names,values)
  
#order factors for legend
  bdf$StateTerritory <- factor(bdf$StateTerritory, 
              levels=c(as.character(topcountries$StateTerritory),"Remaining Average"), 
              labels=legend_labels)
  
#graph  
bdf %>% 
  na.omit() %>% 
  ggplot(aes(Year, Mean_Delegates)) + 
  geom_line(aes(color=StateTerritory)) +
  geom_point() +
  theme_bw() +
  labs(x="Year", y='Mean Delegates', title= 'Non-LDC/SIDS: COP Attendance') +
  theme(plot.title = element_text(hjust = 0.5, face='bold', size = 16),
        axis.title = element_text(size = 14),
        axis.text = element_text(face='bold', size = 12),
        legend.text=element_text(size=10)) +
  scale_color_discrete(name="Country (Average Delegates)")
```

<br> <br>

#### Summary Table: Top 10 & Bottom 10 Attending Non-LDC/SIDS Countries

<br>

```{r, include=TRUE, echo=F, render=lemon_print}
bdf_avg_all <- bdf_avg_all %>% 
  na.omit() %>% 
  mutate(Mean_Delegates = round(Mean_Delegates,1)) %>% 
  arrange(desc(Mean_Delegates)) %>% 
  rename('State Territory' = StateTerritory,'Mean Delegates' = Mean_Delegates)

ht(as.data.frame(bdf_avg_all),10)
```

<br> <br>

^i \<3 tina^

```{r}
# LDC SIDS breakdown
# Add COP locations to last graph
# Address Tina's email
```
